---
title: "HW_9"
author: "Mike Ament"
date: "March 25, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Problem 1: 
Use the code that you worked on in Homework #7 (creating fake data sets) , and re-organize it following the principles of structured programming. Do all the work in a single chunk in your R markdown file, just as if you were writing a single R script. Start with all of your annotated functions, preliminary calls, and global variables. The program body should be only a few lines of code that call the appropriate functions and run them in the correct order. Make sure that the output from one function serves as the input to the next. You can either daisy-chain the functions or write separate lines of code to hold elements in temporary variables and pass the along.

```{r}
## Library

library(ggplot2)

## Preliminary Calls 

treats<- c("ambient", "elevated")  # treatment groups 
n1<- c(20, 20)   # sample sizes 
mean1<- c(50, 60)  ## means 
sd1<- c(10, 10) ## standard deviations 


## Function 1 - get_data

get_data<- function(file_name=NULL) {
  if(is.null(file_name)) {
  df<- data.frame(ID=1:40, 
                  treatment=rep(treats,n), 
                  plant_wt=c(rnorm(n=n1[1], mean=mean1[1], sd=sd1[1]), 
         rnorm(n=n1[2], mean=mean1[2], sd=sd1[2])), 
         plant_ht=c(rnorm(n=n1[1], mean=mean1[1], sd=sd1[1]), 
         rnorm(n=n1[2], mean=mean1[2], sd=sd1[2])))
} else {  ## both need to be on same line 
  df<- read.table(file=file_name,
                  header=TRUE, 
                  sep=",", 
                  stringsAsFactors = FALSE)
  
} 
  return(df)

}


# Function 2 - calc_stuff. In this case, perform the Analysis of Variance. I know that T-tests are used when there are only 2 treatment groups but they are essentially the same.

calc_stuff<- function(x,y) {
  
  AOVmod<- aov(y~x)
  print(AOVmod)
  
} 

# Function 3 - summarize. Summarize the ANOVA table and extract the F-value and p-values. 

summarize<- function(x) {
  z<- summary(x)
  print(z)
  unlist(z)
  AOVsummary<- list(f=unlist(z)[7], p=unlist(z)[9])
  print(AOVsummary)
} 

# Function 4 - graph_results - Graph results as as sideways box and whisker plots

graph_results<- function(file_name) {
  graph1<- ggplot(data=data, aes(x=data[,2], y=data[,3], fill=data[,2]))+
  geom_boxplot() + xlab("Treatment") +ylab("Plant Weight") 
  print(graph1)
}


### String together the functions to generate results 

data<- get_data()
AOV_sum<- calc_stuff(x=data[,2], y=data[,3])
summary_Fp<- summarize(x=AOV_sum)
graph<-graph_results(file_name=data) 






```

Problem 2: 
Once your code is up and working, modify your program do something else: record a new summary variable, code a new statistical analysis, or create a different set of random variables or output graph. Do not rewrite any of your existing functions. Instead, copy them, rename them, and then modify them to do new things. Once your new functions are written, add some more lines of program code, calling a mixture of your previous functions and your new functions to get te job done.

```{r}
### For this problem, I'm going to perform a linear regression of plant weight vs height instead of an ANOVA of plant weight by treatment. I'll use the same global variable Data for this



# Function 2 - calc_stuff   - calcuate the linear regression parameters (slope and intercept) and print them

calc_stuff2<- function(x,y) {
  
  LMmod<- lm(y~x)
  print(LMmod)
  
} 

# Function 3 - summarize  - show the regression analysis table, whcih includes whether the slope and intercept are significantly different from 0. 

summarize2<- function(x) {
  z<- summary(x)
  print(z)
  
} 

# Function 4 - graph_results  - Plant weight vs plant height 

graph_results2<- function(file_name) {
  graph1<- ggplot(data=data, aes(x=data[,3], y=data[,4], fill=data[,2]))+
  geom_point()+ geom_smooth(method = "lm") + xlab("Plant Weight (mg)") +ylab("Plant Height (cm)")
  print(graph1)
}


### String together the functions to generate results 

data<- get_data()
LM_sum<- calc_stuff2(data$plant_wt, data$plant_ht)
summary_LM<- summarize2(LM_sum)
graph<-graph_results2(file_name=data)



```








