---
title: "Homework 4"
author: "Mike Ament"
date: "February 5, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 1: 

Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable zand print the value stored in z.

a) x^a^b
b) (x^a)^b
c) 3x3+2x2+1

```{r}

x<- 1.1
a<- 2.2
b<- 3.3

# Part a
z<- x^a^b
print(z)

# Part b
z<- (x^a)^b
print(z)

# Part c
z<- 3*(x^3) +2*(x^2) +1
print(z)

```

Problem 2: Create the following Vectors using seq and rep
a) (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)
b) (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)
c) (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)

```{r}
# Part a   (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)

a<- seq(1,8,1)
b<- seq(7,1,-1)
z<- c(a,b)
print(z)
# alternatively
z<- c(seq(1,8,1),seq(7,1,-1))
print(z)

# Part b   (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)

x<- seq(1,5,1)
z<- rep(x,x)  # this repeats each element of x, xi number of times
print(z)
  
# Part c  (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)

y<- seq(5,1,-1)
print(y)
z2<- rep(y,x)
print(z2)

```

Problem 3: Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates (If you don’t know what polar coordinates are, read about them on the web or in your calculus textbook).

```{r}
a<- runif(2)
print(a)
x<- a[1]
y<- a[2]

r<- sqrt((x^2)+(y^2))
print(r)

theta<- atan(x/y)
print(theta)

vec<- c(r,theta) ## vector of polar coordinates 


```

Problem 4: 
Suppose that 
``` {r}
q <- c("sheep", "fox", "owl", "ant") 

```
and that queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as

the serpent arrives;
the sheep enters the ark;
the donkey arrives and talks his way to the front of the line;
the serpent gets impatient and leaves;
the owl gets bored and leaves;
the aphid arrives and the ant invites him to cut in line.
Finally, determine the position of the aphid in the line.

```{r}
s<- c("serpent") 
d<- c("donkey") 
a<- c("aphid") 
## add serpent to end of queue  : could also use append to add to the end
q<- c(q,s)
print(q)

## add donkey to front of line 
q<- c(d,q)
print(q)

## subtract the serpent 

q<- q[-6]  ## eliminates the 6th element 
print(q)

## subtract the owl 
q<- q[-4]
print(q)

## insert the aphid between the fox and the ant

q<- c(q[1:3],a,q[4])
print(q)

## alternatively 
z<- append(q, a, after = 3) # first term is vector you're adding too, then the vector you're adding, and the position after which you're adding it
print(z)

## Determine the position of the aphid

aph<- which(q=="aphid")
print(aph) ### will return all the positions where aphid = true

# Extra  - replace the 3rd term with the donkey

q[3]<- d
print(q)

```
Problem 5: Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.

```{r}

x<- seq(1,100,1) ## make the sequence from 1 to 100

g<- x[x%%2!=0 & x%%3!=0 & x%%7!=0] ## extract all the values whose remainders when diveded by 2, 3 and 7 are not equal to 0. 
print(g)

### Alternatively, a much more complicated way 

y<- which(x%%2!=0)
y
z<- which(x%%3!=0)
z
w<- which(x%%7!=0)
w

v<- c(y,z,w)
print(v)

v<- sort(v, decreasing=FALSE)
v

dup<- duplicated(v)
dup
v<- v[!dup]
v
## this actually didn't work so ignore 



```





